---
title: 'ROI data in R'
output:
  rmarkdown::html_vignette:
    toc: true
    toc_depth: 4
    fig_width: 8 
    fig_height: 4 
vignette: >
  %\VignetteIndexEntry{ROI data data in R}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

First we load some packages and set the paths:

```{r warning=FALSE, message=FALSE}
library(ZooDVM)

check.packages <- function(pkg){
  new.pkg <- pkg[!(pkg %in% installed.packages()[, "Package"])]
  if (length(new.pkg))
    install.packages(new.pkg, dependencies = TRUE, INSTALL_opts = '--no-lock')
  sapply(pkg, require, character.only = TRUE)
}
check.packages(c('mgcViz','mgcv','rpart','rpart.plot'))
library(mgcViz)
library(mgcv)
library(rpart)
library(rpart.plot)

datadir <-'C:\\Users\\sven\\Documents\\Zonar\\data\\csv_zonar\\'
roidir <- 'C:\\Users\\sven\\Documents\\Zonar\\data\\csv_zc_roi\\'

```

Then we load Sv, gps and environamnet data.

```{r warning=FALSE, message=FALSE}
#get missions
mm = get_mission(datadir)
sel = grep('*PtSur*',mm$missions)
mission = mm$missions[sel]

#get Sv data
Sv = get_sv(datadir,sel)

#get GPS info
gps = get_gps( ZooDVM::ncpath_from_csvpath(datadir), sel)

#gen environmental data
env = get_env( ZooDVM::ncpath_from_csvpath(datadir), sel)

```

# Getting Zoocam ROIs size class info  

If the standard folder structure is maintained, the pdirectory contianing the roi information can be retrieved through `roipath_from_csvpath()`. ROI information can be read through `get_roi_counts(roidir, mission)` and the ROIS can then be plotted with `plot_roi(rois, roisel, skipdive,...)`:  

```{r getrois, warning=FALSE, message=FALSE}
#get ROIs
#which mission to select?
selroi = grep('*PtSur*',list.files(roipath_from_csvpath(datadir)))
rois = get_roi_counts(roipath_from_csvpath(datadir), selroi)

#plot ROI data
plot_roi(rois,roisel=NULL, skipdive=NULL,dres=1)

```

We could also just plot a selected ROI size:  

```{r roisize, warning=FALSE}
plot_roi(rois, roisel='ROI125')
```

We can also join the acouistic data with the ROIs:  
```{r warning=FALSE, message=FALSE}
#join Sv and ROI data
acroi <- join_ac_roi(Sv, rois)
```


We can have a look at the density distributions of ROI size vs Sv:

```{r}

roivars = names(acroi)[grep('ROI',names(acroi))]

for(v in roivars){
  sub = acroi[acroi[,v]>0,c('Sv200kHz','Sv1000kHz',v)]%>%gather(Freq,Sv,-eval(v))
  p<-ggplot(data=sub)+
    geom_density(aes(Sv, group=Freq, col=Freq))+
    scale_colour_viridis_d(name='')+
    ggtitle(v)+
    theme_classic()+
    theme(legend.position = 'top')
  print(p)
}

```
  
Or we can look at SvDelta vs ROIs:  

```{r}

roivars = names(acroi)[grep('ROI',names(acroi))]

sub = acroi[,c('SvDelta',roivars)]%>%gather(ROI,Count,-SvDelta)
sub = sub[sub$Count>0,]
  p<-ggplot(data=sub)+
    geom_density(aes(SvDelta, group=ROI, col=ROI))+
    scale_colour_viridis_d(name='')+
    ggtitle(v)+
    theme_classic()+
    theme(legend.position = 'top')
  print(p)
  
```


Now we can have a look at joined gams or regression trees for example to better understand the relationship between the ROIs and the acoustic data.  

THis is not necessarily the best example but it is an analysis for each ROI size class during the day, with a smoother function for Sv 200 kHz, Sv 1000 kHz and an interaction term:  

```{r message=FALSE,results="asis", warning=FALSE}
#make gams and plots

#set depth cut value
dcut=200
#select daytime to analyze:
daytime = 'Day' 

#loop through the different ROIs
for(v in c('ROI25','ROI45','ROI75','ROI125','ROI200')){
    cat("  ###",  v," for ", daytime,"  ")

  g1 =mgcv::gam(data=acroi[acroi$Depth<dcut & acroi$sun==daytime,],eval(parse(text=v))~s(Sv200kHz)+s(Sv1000kHz)+s(Sv200kHz, Sv1000kHz))
  summary(g1)
  
  b <- getViz(g1)
  print(plot(b, allTerms = T) +  l_fitRaster() + l_fitContour() + l_points(alpha=0.2)+ l_fitLine(linetype = 1, size=1.2) + l_fitContour() + 
          l_ciLine(colour = 2) +
          scale_fill_gradientn(colors=rev(pals::brewer.rdylbu(12)) , na.value='transparent',oob=scales::squish)+theme_classic(), pages = 1)
}

```

Or we look at it from another perspective:  

```{r message=FALSE,results="asis", warning=FALSE}
#make gams and plots

#set depth cut value
dcut=200
#select daytime to analyze:
daytime = 'Day' 
rvars =c('ROI25','ROI45','ROI75','ROI125','ROI200')
#loop through the different ROIs
for(v in c('Sv200kHz','Sv1000kHz','SvDelta')){
    cat("  ###",  v," for ", daytime,"  ")

  g1 =mgcv::gam(data=acroi[acroi$Depth<dcut & acroi$sun==daytime,],eval(parse(text=v))~s(ROI25)+s(ROI200)+s(ROI25, ROI200)+s(ROI25, ROI125)+s(ROI25, ROI75)+s(ROI75, ROI200))
  summary(g1)
  
  b <- getViz(g1)
  print(plot(b, allTerms = T) +  l_fitRaster() + l_fitContour() + l_points(alpha=0.2)+ l_fitLine(linetype = 1, size=1.2) + l_fitContour() + 
          l_ciLine(colour = 2) +
          scale_fill_gradientn(colors=rev(pals::brewer.rdylbu(12)) , na.value='transparent',oob=scales::squish)+theme_classic(), pages = 1)
}

```
Similiarly we can look at regression trees to get a best classifier. Here an example for ROI 25:  


```{r}

m1 <- rpart(
  formula = (ROI25) ~ Sv1000kHz + Sv200kHz + SvDelta,
  data    = acroi)
m1$variable.importance
rpart.plot(m1,type=1)
```
  
Or for ROI 200:
    
```{r}


m1 <- rpart(
  formula = (ROI200) ~ Sv1000kHz + Sv200kHz + SvDelta,
  data    = acroi)
m1$variable.importance
rpart.plot(m1,type=1)
```
